cvx-flats

### 2020-04-13 neurips-20/cvx-flats.tex
Taking a pass on theorem 1 and org

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\elic}{\mathrm{elic}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{ccdim}}

@@

\author{%
	Peter Bartlett\\
	UC Berkeley\\
	\texttt{bartlett@cs.berkeley.edu}
	\And
  Jessie Finocchiaro\\
%  Department of Computer Science\\
  CU Boulder\\
  \texttt{jefi8453@colorado.edu} 
  % examples of more authors
  \And
   Rafael Frongillo\\
%   Department of Computer Science\\
   CU Boulder\\
  % Address \\
   \texttt{raf@colorado.edu} 
   \And
   Bo Waggoner\\
%   Department of Computer Science\\
   CU Boulder \\
  % Address \\
   \texttt{bwag@colorado.edu} 
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

\cite{finocchiaro2019embedding} presents the notion of a surrogate loss \emph{embedding} a discrete loss-- typically by a polyhedral (piecewise linear and convex) surrogate.
For all discrete losses, $\eliccvx(\ell) \leq \elicpoly(\ell) \leq \elicembed(\ell)$.
However, one open question that remains from their work is if these bounds are equal. \jessie{\ldots} \jessie{MOVE}

@@

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given an elicitable property $\gamma$.
	Fix $p \in \simplex$, and let $r$ be a report such that $p \in \gamma_r$.  
	If there is not $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals_+$ that indirectly elicits $\gamma$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	We prove the contrapositive; if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

	Suppose we have $L:\reals^d \times \Y \to \reals$ eliciting $\Gamma$ and link $\psi : \reals^d \to \R$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Fix any $p \in \simplex$, and take any $r \in \gamma(p)$.
	We argue that there must be some $u \in \reals^d \in \Gamma(p)$ so that $r = \psi(u)$, since $\Gamma(p) = \emptyset$ otherwise.
	We know that $r \in \gamma(p) \iff u \in \Gamma(p) \iff \vec 0 \in \partial L(u; p) \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the nonnegative weighted Minkowski sum by convex analysis results \jessie{And mild assumptions on the loss?}.

	The weighted Minkowski sum is in $d$-dimensional space as $L$ is defined on $\reals^d$.
	We claim there exists a matrix $W \in \reals^{d+1 \times n}$ defining the flat $F = \{q \in \reals^n : Wq = \vec 0 \}$; $F$ is then $(n-(d+1))$-dimensional by the rank-nullity theorem, so we just need to show such a $W$ also keeps $F \cap \simplex \subseteq \gamma_r$.
	As $\Gamma_u \subseteq \gamma_r$, it is sufficient to show $F \cap \simplex \subset \Gamma_u$.

	Construct $W$ by taking each column to describe $\partial L(u, y)$ for each $y \in \Y$.
	Since $\partial L(u, y)$ is $d$-dimensional, the affine subspace containing each subgradient can be described with $d+1$ constraints. \jessie{??}
	We can see that $p \in F$ since $\vec 0 \in \oplus_y p_y \partial L(u, y) \implies Wp = \vec 0$ by construction of $W$.
	Moreover, if there was a $\hat q \in \simplex \setminus \gamma_r$ such that $W \hat q = \vec 0$, then we would contradict the indirect elicitation of $\gamma$ since we would not have $\Gamma_{u} \subseteq \gamma_{\psi(u)}$ since we would have $\hat q \in \Gamma_u$ but $\hat q \not \in \gamma_{\psi(u)}$.
	Therefore,  we have an $(n - d - 1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$ for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{proof}

@@

Theorem~\ref{thm:calib-iff-consistent} allows us to now relate convex elicitation complexity and convex calibration dimension introduced by~\cite[Definition 10]{ramaswamy2016convex}.
\begin{proposition}
	Consider the discrete loss $\ell : \R \times \Y \to \reals_+$ and $\gamma:= \prop{\ell}$.
	Then $\eliccvx(\gamma) \leq \ccdim(\ell)$.
	\jessie{Move this to section 5?}
	\jessie{Need to define $\eliccvx$ and $\ccdim$.}
\end{proposition}
\begin{proof}
	Let $d$ be the convex calibration dimension of $\ell$, and $(L, \psi)$ be calibrated with respect to $\ell$ with $L$ convex.
	Then $(L, \psi)$ is also consistent with respect to $\ell$, and therefore $L$ indirectly elicits $\gamma$.
\end{proof}

@@

\jessie{Abstain example... maybe... hopefully...}

>>>>

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\elic}{\mathrm{elic}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{ccdim}}
\newcommand{\codim}{\mathrm{codim}}

@@

\author{%
  Jessie Finocchiaro\\
%  Department of Computer Science\\
  CU Boulder\\
  \texttt{jefi8453@colorado.edu} 
  % examples of more authors
  \And
   Rafael Frongillo\\
%   Department of Computer Science\\
   CU Boulder\\
  % Address \\
   \texttt{raf@colorado.edu} 
   \And
   Bo Waggoner\\
%   Department of Computer Science\\
   CU Boulder \\
  % Address \\
   \texttt{bwag@colorado.edu} 
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

@@

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a finite elicitable property $\gamma$.
	Fix $p \in \simplex$, and let $r$ be a report such that $p \in \gamma_r$.  
	If there is no $(n - d - 1)$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals_+$ that indirectly elicits $\gamma$.
	Moreover, this must hold for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{theorem}
\begin{proof}[Proof intuition]
	\jessie{BTW: $n$, or $\|p\|_0$?}
	We prove the contrapositive; if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

	Suppose we have $L:\reals^d \times \Y \to \reals$ eliciting $\Gamma$ and link $\psi : \reals^d \to \R$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, i.e. $L$ indirectly elicits $\gamma$.
	Fix any $p \in \simplex$, and take any $r \in \gamma(p)$.
	We argue that there must be some $u \in \Gamma(p) \cap \reals^d$ such that $r = \psi(u)$, since $\Gamma(p) = \emptyset$ otherwise.
	We know that $r \in \gamma(p) \iff u \in \Gamma(p) \iff \vec 0 \in \partial L(u; p) \iff \vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the nonnegative weighted Minkowski sum by convex analysis results \jessie{And mild assumptions on the loss?}.

	Observe that the weighted Minkowski sum is in $d$-dimensional space as $L$ is defined on $\reals^d$.
	As $\partial L(u;p)$ is defined in $n$-dimensional space over $\simplex$, the flats of $L(u;p)$ are at least $(n-d-1)$-dimensional as the subgradient is in $d$-dimensional space, and the affineness of being a flat allows us to remove one additional dimension.
	Let $F$ be an $(n-d-1)$-dimensional flat to $L(u;p)$.
	For all $q \in F$, we must have $u \in \Gamma(q)$ since \jessie{JUSTIFICATION} there is a bijection from $F \in \partial L(u;p)$ to $\Gamma(p)$. Therefore, $q \in F \implies u \in \Gamma(q)$ as well.
	Therefore, we have $F \cap \simplex \subseteq \Gamma_u \subseteq \gamma_r$, where the latter follows from indirect elicitation.

%	We claim there exists a matrix $W \in \reals^{d+1 \times n}$ defining the flat $F = \{q \in \reals^n : Wq = \vec 0 \}$; $F$ is then $(n-(d+1))$-dimensional by the rank-nullity theorem, so we just need to show such a $W$ also keeps $F \cap \simplex \subseteq \gamma_r$.
%	As $\Gamma_u \subseteq \gamma_r$, it is sufficient to show $F \cap \simplex \subset \Gamma_u$.
%	 
%	Construct $W$ by taking each column to describe $\partial L(u, y)$ for each $y \in \Y$.
%	Since $\partial L(u, y)$ is $d$-dimensional, the affine subspace containing each subgradient can be described with $d+1$ constraints. \jessie{??}
%	We can see that $p \in F$ since $\vec 0 \in \oplus_y p_y \partial L(u, y) \implies Wp = \vec 0$ by construction of $W$.
%	Moreover, if there was a $\hat q \in \simplex \setminus \gamma_r$ such that $W \hat q = \vec 0$, then we would contradict the indirect elicitation of $\gamma$ since we would not have $\Gamma_{u} \subseteq \gamma_{\psi(u)}$ since we would have $\hat q \in \Gamma_u$ but $\hat q \not \in \gamma_{\psi(u)}$.
%	Therefore,  we have an $(n - d - 1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$ for all $p \in \simplex$ and $r \in \gamma(p)$.
\end{proof}

\begin{theorem}[Generalization to infinite outcomes?]
	Suppose we are given an elicitable property $\gamma$.
	Fix $p \in \simplex$, and let $r$ be a report such that $p \in \gamma_r$.  
	If there is no $\codim(\partial L(r;p))-1$-dimensional flat $F$ such that $F \subseteq \gamma_r$ and $p \in F$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
	Moreover, this must hold for all $p \in \mathcal{P}$ and $r \in \gamma(p)$.
\end{theorem}

@@

Theorem~\ref{thm:calib-iff-consistent} allows us to now relate convex elicitation complexity and convex calibration dimension introduced by~\cite[Definition 10]{ramaswamy2016convex}.
\begin{proposition}
	Consider the discrete loss $\ell : \R \times \Y \to \reals_+$ and $\gamma:= \prop{\ell}$.
	Then $\eliccvx(\gamma) \leq \ccdim(\ell)$.
	\jessie{Need to define $\eliccvx$ and $\ccdim$.}
\end{proposition}
\begin{proof}
	Let $d$ be the convex calibration dimension of $\ell$, and $(L, \psi)$ be calibrated with respect to $\ell$ with $L$ convex.
	Then $(L, \psi)$ is also consistent with respect to $\ell$, and therefore $L$ indirectly elicits $\gamma$.
\end{proof}

@@

\cite{finocchiaro2019embedding} presents the notion of a surrogate loss \emph{embedding} a discrete loss-- typically by a polyhedral (piecewise linear and convex) surrogate.
If $\gamma$ is the finite property elicited by the discrete loss $\ell$, then for all discrete losses, $\eliccvx(\gamma) \leq \elicpoly(\gamma) \leq \elicembed(\ell)$.
However, one open question that remains from their work is if these bounds are equal. \jessie{\ldots}

\subsection{Example: Abstain loss}
\jessie{Abstain example... maybe... hopefully...}
